---
title: "Daily Assignment 5: Madi Schartz"
Author: Madi Schartz
format: html
execute: 
  echo: true
---
###Questions:

###question: 1. Examine at the dataset using the ?Help page

```{r}
library(palmerpenguins)
data("penguins")
?penguins

```

###question: 2. what is the class of the penguins data set?
answer: The data is in a table format and data frame.
```{r}
class(penguins)
```

###question:  3. what is the structure of the penguins data set?
answer: The structure of the penguins data set is in a table where there are 8 variables recorded consisting of: species type, island, bill length in mm, bill depth in mm, flipper length in mm, body mass in grams, the sex of the penguin, and the year.  

```{r}
str(penguins)
```

###question: 4. what are the dimensions of the penguins dataset?
answer: the dimensions of this data set are 8 columns and 344 rows of data recorded. 
```{r}
dim(penguins)
```

###question: 5. what are the column names of the penguins dataset?
answer: The column names consist of: species type, island, bill length in mm, bill depth in mm, flipper length in mm, body mass in grams, the sex of the penguin, and the year. 
```{r}
colnames(penguins)
```

### question: 6. what type of data is `flipper_length_mm` and `Island`?
answer: The flipper length is an example of integer data and the island is an example of factor data.
```{r}
class(penguins$flipper_length_mm) #Numeric
 class(penguins$island) # Factor (Categorical)
```

###question: 7. what is the mean flipper length of the penguins?
answer: The mean flipper length is 200.9152 mm
```{r}
mean(penguins$flipper_length_mm, na.rm = TRUE)

```

###question:  8. what is the standard deviation of flipper length in the penguins?
answer: The standard deviation of the flipper length in the penguins is 14.06171 units
```{r}
 sd(penguins$flipper_length_mm, na.rm = TRUE)
```

###question:  9. what is the median body mass of the penguins?
answer: the median body mass of the penguins is 4050 g
```{r}
median(penguins$body_mass_g, na.rm = TRUE)
```

###question:  10. what is the Island of the 100th penguin?
answer: The island of the 100th penguin is Biscoe Dream Torgersen.
```{r}
penguins$island[100]
```


